---
title: "Visualizing-what-convnets-learn"
author: "Etienne Rolland"
date: "04/05/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Visualizing-what-convnets-learn

Most of the code came from this page came from this [chapter of the book Deep learning with R](https://jjallaire.github.io/deep-learning-with-r-notebooks/notebooks/5.4-visualizing-what-convnets-learn.nb.html
). The goal is to check what the differents conv_nets has learned before submitting. At the moment I can't run it since the resnet has been built using the sequential API.

## Set up

```{r}
library(tidyverse)
library(keras)
library(tensorflow)
library(reticulate)
```


```{r}
use_python("/usr/bin/python3.5", required = TRUE)
tf <- tensorflow::tf
sess <- tf$Session()
```

```{r}
image_path<-'plant-pathology-2020-fgvc7/images/'
```

```{r}
#reticulate::virtualenv_install(packages="pandas") 
```

```{r}
keras::use_implementation("keras")
keras::use_backend("tensorflow")

```

```{r}
reticulate::py_config()
```


## Loading model


```{r}
conv_base <- application_resnet50(weights = 'imagenet', include_top = FALSE, input_shape = c(448, 448, 3))
```

```{r}
unfreeze_weights(conv_base, from="res5a_branch2a")
```


```{r}
model <- keras_model_sequential() %>% 
        conv_base %>% 
        layer_global_max_pooling_2d() %>% 
        layer_batch_normalization() %>%
        layer_dropout(rate=0.5) %>%
        layer_dense(units=4, activation="softmax")
```

```{r}
model
```

```{r}
checkpoint_dir <- "fine_tuned_models/"

model %>% load_model_weights_hdf5(
  file.path(checkpoint_dir,"Fine_tuned_Resnet50_res5a_05.hdf5")
)
```

# Visualization


```{r}
img_path<-"plant-pathology-2020-fgvc7/images/Train_1.jpg"

```


NB : it is really important to use imagenet_preprocess_net(), otherwise the predictions does not reflect what the model has learned.

```{r}
img <- image_load(img_path, target_size = c(448, 448)) %>% 
  # Array of shape (224, 224, 3)
  image_to_array() %>% 
  # Adds a dimension to transform the array into a batch of size (1, 224, 224, 3)
  array_reshape(dim = c(1, 448, 448, 3)) %>% 
  # Preprocesses the batch (this does channel-wise color normalization)
  imagenet_preprocess_input()
```


```{r}
preds <- model %>% predict(img)
print(c("healthy", "multiple_diseases", "rust", "scab"))
print(preds)
```

```{r}
summary(model)
```



























